Well-Designed Environmental Markets Enable Large-Scale Marine Conservation

### 2019-10-26 docs/Markets_Conservation_revised_Oct24.tex
Updating with John's changes

Recognizing the need to protect marine biodiversity and ecosystem services, various international bodies have committed to dramatically expand marine protection around the world by protecting 30\% of the oceans \cite{oleary_2016,dinerstein_2019}, and that large portions of these will be entirely no-take \cite{sala_2017,sala_2018}. While most marine protected areas (MPAs) implemented to date have been small, it is widely-recognized that to meet these goals, very large MPAs must also be part of the strategy. But would any country rationally close 30\%, 80\% or even 100\% of its waters to fishing if this means losing all fishing revenue from within the closed area? At first glance, the answer is probably ``no''. Losing this important source of revenue would cripple many fishing-dependent economies, particularly the Pacific Island nations who are viewed by many as viable candidates for large closures \cite{mcleod_2019}. On the other hand, the closure may generate substantial ``spillover'' of larvae and adult fish \cite{ramesh_2019,hernndez_2019} and other benefits into adjacent waters which could, in principle, offset these losses. The problem with this argument is that for very large MPAs, these spillover benefits could accrue to other nations and the high seas \cite{agardy_2018}, with no obvious mechanism for the conserving country to recoup them. A viable solution may lie in international fishing effort markets, where nations trade the right to fish. The design and effectivness of fishing effort markets for fisheries management have been explored before \cite{havice_2010,havice_2013}, but little attention has been given to the role these markets can play in conservation. Here, we will show how fishing effort markets can be designed (if new) or modified (if already existing) to incentivize implementation of large-scale marine protected areas.

We are motivated by a real-world but relatively understudied institution called the Parties to the Nauru Agreement (PNA). The PNA is a coalition of nine Pacific island nations that collectively manages tuna purse seine fishing in its members' waters (Fig. \ref{fig:PNA_map}) \cite{havice_2013,aqorau_2018}. These waters account for 14.5 million km\textsuperscript{2} (an area four times larger than the continental US), and over 60\% of skipjack tuna catch in the Western Central Pacific \cite{havice_2013}. The PNA manages this tuna purse seine fishery using a vessel-day scheme (VDS) where total annual fishing effort is capped at around 45,000 vessel-days. Vessel-days are allocated across the nine nations, which then lease them to (mostly foreign) fishing vessels. A vessel-day grants a fishing vessel the right to fish for 24 hours within one of the nine Exclusive Economic Zones (EEZs) within the PNA. Member nations derive enormous benefits from leasing these fishing rights to foreign fleets, in some cases exceeding half of a country's GDP (XXX). In addition to highly productive tuna fisheries, the PNA waters provide a wealth of ecosystem benefits, hence the focus on large-scale conservation efforts in the region \cite{mcleod_2019}. In 2015, Kiribati, a PNA member, created the Phoenix Islands Protected Area, one of the largest protected areas on earth (408,250 km\textsuperscript{2}), and Palau will close 80\% of its national waters to fishing by December 2020. We draw from the PNA's market-based approach to managing tuna fisheries and build on it to show how the market can be designed to incentivize conservation.

Not all market-based approaches to environmental management contain the same conservation incentives. A pervasive finding across a range of natural resources is that features of markets, such as the allocation of rights, can have implications for the market's functioning \cite{libecap_1989}. In the context of fishing effort markets, we find that two market design features are pivotal in determining the incentives for large-scale marine conservation: trading and allocation rules. But why would these design features of a fisheries market affect the incentives for conservation? Consider the incentives for a country to close 100\% of its waters. Such a closure would surely benefit other countries through the spillover of fish from the protected area to the waters of neighboring countries. If the conserving country could lease the rights to that spillover to adjacent countries, this could offset the foregone fisheries revenue. But if the conserving country was not allowed to lease these rights, then it would lose all of its fishing revenue. The rules for how fishing rights are allocated across countries are equally important. Suppose rights are allocated each year based on the previous year's fishing effort: the more a country fishes, the more it gets allocated the next year. This would clearly disadvantage a conservation-minded country and, in fact, might reward undesirable behavior. In the following section, we will show how trading and allocation rules can shape the incentives that may facilitate or hinder large-scale conservation. We will then make use of vessel-tracking data to analyze a real-world case of large-scale marine conservation under fishing effort markets.

@@

We first simulate a fishery where trading between countries is not allowed. This represents the status quo of any nation unilaterally engaging in large-scale conservation. Intuitively, we find that a spatial closure in Country 1 will always result in a loss in revenues to Country 1 (Fig. \ref{fig:PNA_model}A). Higher within-country stock movement ($\theta$ = 0 implies no within-country movement and $\theta = 1$ implies that fish are well-mixed within the fishing season) allows vessels to harvest the stock within the remaining open area, lowering the cost to Country 1. The closure-to-cost ratio for any reserve size is greater than 1:1 (XXX Why) when stock movement is low (\emph{i.e.} $\theta < 0.2$), implying that a 30\% closure would result in at least a 30\% loss in revenues. Even for a highly mobile stock where fish can move in and out of the reserve, a spatial closure reduces the amount of biomass that is available for harvest in the conserving country (\emph{i.e.} biomass outside the reserve), which reduces vessels' willingness to pay to fish in such waters (Fig. S1). When countries cannot trade, the costs of conservation are incurred by Country 1, but the benefits are received by the eight other countries (revenues increase between 0\% and 6\% each; Fig. S2) and the high seas. This benchmark simulation highlights the misalignment of incentives, where a conservation-minded country incurs large costs and provides public benefits to other nations with no mechanism for recouping these benefits.

How would trading between countries change these results? We simulate the same fishery, but now allow for vessel-days to be traded across countries (as is the case for the PNA). As before, a closure in Country 1 lowers the value of vessel-days in that country (because the fishable water shrinks). But increased biomass in other countries causes their vessel day prices to increase. As a result, vessel-days from Country 1 are traded to Countries 2 to 9, until prices equalize across countries (Fig. S3). Under this market design, revenue losses to Country 1 are less than 1\%, compared to the base case with no reserve (Fig. \ref{fig:PNA_model}B; note that the vertical axis now only ranges from 0 to 0.8\% instead of from 0 to 100\% as in Fig. \ref{fig:PNA_model}A). With trading, the relative revenue drop will always be smaller than the relative effort drop, and the opposite is observed when there is no trading (Fig. S4). Overall, this shows that 88\% to 99\% of the costs of conservation can be avoided if markets are designed to allow trading (Fig. \ref{fig:PNA_model}C).

@@

A large-scale MPA was recently implemented in PNA waters, providing the ideal empirical setting to test our predictions. In January 2015, Kiribati closed 11.5\% of its EEZ to implement the Phoenix Island Protected Area (PIPA), effectively displacing all fishing effort within its boundaries\cite{mccauley_2016,mcdermott_2018}. By protecting important tuna spawning habitat, PIPA may provide important recruitment and biomass benefits to adjacent waters \cite{hernndez_2019}. We combine vessel-tracking data \cite{kroodsma_2018} and country-level license revenue data reported by the Pacific Islands Forum Fisheries Agency (FFA) \cite{ffa_2017} to quantify the displacement of vessel-days and the likely costs of conservation. Of the 318 tuna purse seine vessels that fished in PNA waters between 2012 and 2018, 64 ``displaced'' vessels fished within PIPA at least once prior to its implementation and 254 ``non-displaced vessels'' never fished in PIPA waters but fished within the PNA waters. We use the vessel-tracking data to calculate vessel-days in the same way that the PNA does (see Methods for details) and the redistribution of fishing events before and after the implementation of PIPA.

Consistent with our model's prediction, after PIPA was implemented, displaced vessels relocated largely outside of Kiribati, and into other PNA countries' waters (Fig \ref{fig:fishing_raster_diff} and Figs. S6). Indeed, from 2014 to 2015, displaced vessels spent 2,115 fewer vessel-days (a 25\% reduction) in Kiribati, and 2,298 fewer vessel-days in PNA waters (a 17 \% reduction; Fig. \ref{fig:empirical}). On the other hand, non-displaced vessels spent 4,656 additional days in Kiribati during 2015, and an additional 9,598 vessel-days in PNA waters. By 2018, we observe a net decrease of vessel-days within Kiribati, from 12,671 in 2014 to just 7,677 in 2018, with displaced vessels driving the decrease (Fig. \ref{fig:empirical}A). However, aggregate effort at the PNA-level remains relatively constant and we do not observe a ``fishing the line'' effect (Fig S7). The reduction in effort to Kiribati and constant effort at the PNA-level suggest that trading facilitated redistribution of effort within PNA waters, just as the model predicted.

The decrease of vessel-days in Kiribati could be alternatively (or jointly) explained by changes in oceanographic conditions that drive the distribution of target species and fishing vessels. For example, as El Ni\~no events develop, tuna species are known to shift longitudinally across PNA waters, causing vessels to redistribute \cite{aqorau_2018,hanich2018unraveling}. However, the aggregate decrease that we observe for Kiribati in Fig. \ref{fig:empirical} is driven by the relocation of vessels that historically fished within PIPA, not by the entire fleet. Oceanographic conditions should be influencing the entire fleet whereas the closure of PIPA should have a stronger impact on vessels that used to fish there. Thus, the large drop in effort within Kiribati for the displaced vessels (relative to the non-displaced vessels) is consistent with the argument that PIPA displaced these vessels to waters outside of Kiribati. One could argue that differences in vessel characteristics between displaced and non-displaced vessels may influence their ability to redistribute or take advantage of different oceanographic conditions in Kiribati. It should be noted that, on average, displaced vessels have smaller crew sizes, more engine power, and are larger than non-displaced vessels (Table S2, Fig. S5). Another alternative explanation is a change in international relations between Kiribati (or other PNA countries) and the flag country of a particular vessel. Ruling out alternative explanations for observed effects is an important component of careful empirical evaluation of MPAs \cite{ferraro2018causal}: as such, our results should not be interpreted as the direct causal impacts of PIPA but are better viewed as patterns that are consistent with the predictions of our theoretical model.

@@

%%%% PLEASE TAKE A LOOK AT THE SECOND SENTENCE OF THIS PARAGRAPH %%%%
While trading would allow Palau, Kiribati, and other PNA members to reduce revenue losses from large-scale conservation, our model indicates that moving from a 40\% biomass-based allocation rule to a 100\% biomass-based rule would ensure long-term financial security in the presence of large-scale MPAs, and further incentivize conservation within the PNA. This is a model prediction that cannot be empirically tested in the PNA because the rule has not been experimentally modified. The PNA countries have shown that rights-based management of transboundary resources can result in large management and economic benefits \cite{havice_2013,aqorau_2018}. By facilitating trade and allocating rights based on biomass, they may also have become pioneers in effective large-scale marine conservation in a market-based setting.

There are, however, a series of important considerations to take into account. First, in our model only one of the countries considers the implementation of a protected area. An interesting extension could consider cooperative conservation. In that setting, a group of countries could coordinate on an optimized, large-scale strategic closure -perhaps using vessel-day trading as a means of compensation- in a manner similar to the model above. Future research could explore the role of heterogeneous costs and benefits between countries, and the way in which these can shape conservation outcomes in a market-based setting. A second consideration is the role of the high seas: environmental markets require secure property rights, which are lacking in the high seas \cite{crespo_2019}. Benefits accruing to the high seas could potentially be eroded by the prevailing open access conditions. Therefore, a conservation-minded nation has no mechanism to capture the benefits provided to the high seas, highligting the importance of empowering high seas governance and transboundary cooperation \cite{aqorau_2018,crespo_2019}. Finally, our work focuses on fisheries and large-scale conservation, but the framework could potentially be expanded and applied to other systems and natural resources. For example, similar mechanisms could be implemented in markets for tradable water rights \cite{rosegrant_1994} or game hunting \cite{hutton_2003} where secure property rights and the presence of a market may incentivise users to conserve the resource in question.

@@

Automatic Identification Systems (AIS) are on-board devices that provide at-sea safety and prevent ship collisions by broadcasting vessel position, course, and activity to surrounding vessels. These broadcast messages can be received by satellites and land-based antennas. We use AIS data provided by Global Fishing Watch \cite{kroodsma_2018} to track 318 tuna purse seiners that fished within the PNA. For every georeferenced position, we observe the time spent (defined as the time since the last position), and whether the vessel was activly fishing versus only transiting. Of the 318 tuna purse seine vessels that fished in PNA waters between 2012 and 2018, 64 ``displaced'' vessels fished within PIPA at least once prior to its implementation, the remianing 254 ``non-displaced vessels'' never fished in PIPA waters. Our dataset contains more than 37 million geo-referenced positions for these 318 tuna purse seiners. We use these data to calculate vessel-days (the metric used by the PNA), and to track the spatial redistribution of displaced vessels. A comparision of vessel characteristics between displaced and non-displaced vessels is presented in Table S2 and Figure S5.

We use these data to calculate the number of vessel-days that the 318 purse seiners spent in each PNA country and in PNA waters as a whole (Fig. \ref{fig:empirical}). The vessel-day equivalent of a day of fishing depends on vessel size, a measure used to control for effort creep. A day of activity by vessels smaller than 50 meters long counts as half a vessel-day, a day of activity by vessels 50 - 80 meters counts as one day, and a day of activity by vessels larger than 80 meters counts as 1.5 vessel-days. Vessel length is an observable characteristic in our dataset, and therefore vessel-days calculated in our analyses correspond to the PNA definition of vessel-days.

@@

We begin by filtering the data to keep only positions labeled as fishing events. We then create a gridded version of the data for each year and group (\emph{i.e.} displaced and non-displaced) by binning the coordinates to a 1-degree grid and summing all fishing hours for a given grid cell, resulting in 14 (7 years for two groups) gridded datasets of fishing hours. For each group of vessels, we then calculate the average fishing hours before (2012 - 2014, inclusive) and after (2015 - 2018, inclusive) the implementation of PIPA, resulting in 4 datasets of mean fishing effort (before and after for displaced and non-displaced vessels). We then calculate the change in effort allocation between these two periods (after minus before) for each group (Fig. \ref{fig:fishing_raster_diff}A-B). The redistribution by non-displaced vessels (Fig. \ref{fig:fishing_raster_diff}B) therefore provides a baseline of redistribution. We then compare the changes of displaced vessels to those by non-displaced vessels (Fig. \ref{fig:fishing_raster_diff}C). The spatial redistribution patterns of displaced vessels relative to non-displaced vessels suggest that some relocated to other waters in Kiribati (\emph{i.e.} Gilbert islands and Line islands), but also the Marshall Islands, Tuvalu, Nauru, and the high seas.

\subsubsection{Revenues and catches}

>>>>

Recognizing the need to protect marine biodiversity and ecosystem services, various international bodies have committed to dramatically expand marine protection around the world by protecting 30\% of the oceans \cite{oleary_2016,dinerstein_2019}, with a focus on no-take Marine Protected Areas (MPAs) \cite{sala_2017,sala_2018}. While most MPAs implemented to date have been small, it is widely-recognized that, to meet these goals, very large MPAs must also be part of the strategy. But would any country rationally close 30\%, 80\% or even 100\% of its waters to fishing if this means losing all fishing revenue from within the closed area? At first glance, the answer is probably ``no''. Losing this important source of revenue would cripple many fishing-dependent economies, particularly the Pacific Island nations who are viewed by many as viable candidates for large closures \cite{mcleod_2019}. On the other hand, the closure may generate substantial ``spillover'' of larvae and adult fish \cite{ramesh_2019,hernndez_2019} and other benefits into adjacent waters which could, in principle, offset these losses. The problem with this argument is that for very large MPAs, these spillover benefits could accrue to other nations and the high seas \cite{agardy_2018}, with no obvious mechanism for the conserving country to recoup them. A viable solution may lie in international fishing effort markets, where nations trade the right to fish across international boundaries. The design and effectiveness of fishing effort markets for fisheries management have been explored before \cite{havice_2010,havice_2013}, but little attention has been given to the role these markets can play in conservation. Here, we will show how fishing effort markets can be designed (if new) or modified (if already existing) to incentivize implementation of large-scale marine protected areas.

We are motivated by a real-world but relatively understudied institution called the Parties to the Nauru Agreement (PNA). The PNA is a coalition of nine Pacific island nations that collectively manages tuna purse seine fishing in its members' waters (Fig. \ref{fig:PNA_map}) \cite{havice_2013,aqorau_2018}. These waters account for 14.5 million km\textsuperscript{2} (an area four times larger than the continental US), and over 60\% of skipjack tuna catch in the Western Central Pacific \cite{havice_2013}. The PNA manages this tuna purse seine fishery using a vessel-day scheme (VDS) where total annual fishing effort is capped at around 45,000 vessel-days. Vessel-days are allocated across the nine nations, which then lease them to (mostly foreign) fishing vessels. A vessel-day grants a fishing vessel the right to fish for 24 hours within one of the nine Exclusive Economic Zones (EEZs) within the PNA. Member nations derive enormous benefits from leasing these fishing rights to foreign fleets, in some cases exceeding half of a country's GDP (XXX). In addition to highly productive tuna fisheries, the PNA waters provide a wealth of ecosystem benefits, hence the focus on large-scale conservation efforts in the region \cite{mcleod_2019}. In 2015, Kiribati, a PNA member, created the Phoenix Islands Protected Area, one of the largest protected areas on earth (408,250 km\textsuperscript{2}), and Palau will close 80\% of its national waters to fishing by December 2020. We draw from the PNA's market-based approach to managing tuna fisheries and build on it to show how a market can be designed to incentivize conservation.

Not all market-based approaches to environmental management contain the same conservation incentives. A pervasive finding across a range of natural resources is that features of markets, such as the allocation of rights, can have implications for the market's functioning \cite{libecap_1989}. In the context of fishing effort markets, we find that two market design features are pivotal in determining the incentives for large-scale marine conservation: trading and allocation rules. But why would these design features of a fisheries market affect the incentives for conservation? Consider the incentives for a country to close 100\% of its waters. Such a closure would surely benefit other countries through the spillover of fish from the protected area to the waters of neighboring countries. If the conserving country could trade the rights to that spillover to adjacent countries, this could offset the foregone fisheries revenue. But if the conserving country was not allowed to trade these rights, then it would lose all of its fishing revenue. The rules for how fishing rights are allocated across countries are equally important. Suppose rights are allocated each year based on the previous year's fishing effort: the more a country fishes, the more it gets allocated the next year. This would clearly disadvantage a conservation-minded country and, in fact, might reward undesirable behavior. In the following section, we will show how trading and allocation rules can shape the incentives that may facilitate or hinder large-scale conservation. We will then make use of vessel-tracking data to analyze a real-world case of large-scale marine conservation under fishing effort markets.

@@

We first simulate a fishery where trading between countries is not allowed. This represents the status quo of any nation unilaterally engaging in large-scale conservation. Intuitively, we find that a spatial closure in Country 1 will always result in a loss in revenues to Country 1 (Fig. \ref{fig:PNA_model}A). Higher within-country stock movement ($\theta$ = 0 implies no within-country movement and $\theta = 1$ implies that fish are well-mixed within the fishing season) allows vessels to harvest the stock within the remaining open area, lowering the cost to Country 1. The closure-to-cost ratio for any reserve size is greater than 1:1 (XXX Why) when stock movement is low (\emph{i.e.} $\theta < 0.2$), implying that a 30\% closure would result in at least a 30\% loss in revenues. XXXX The intuition for why the ratio is higher than 1:1 is that fishing cost is a convex function of reserve size. XXXX Even for a highly mobile stock where fish can move in and out of the reserve, a spatial closure reduces the amount of biomass that is available for harvest in the conserving country (\emph{i.e.} biomass outside the reserve), which reduces vessels' willingness to pay to fish in such waters (Fig. S1). When countries cannot trade, the costs of conservation are incurred by Country 1, but the benefits are received by the eight other countries (revenues increase between 0\% and 6\% each; Fig. S2) and the high seas. This benchmark simulation highlights the misalignment of incentives, where a conservation-minded country incurs large costs and provides public benefits to other nations with no mechanism for recouping these benefits.

How would trading between countries change these results? We simulate the same fishery, but now allow for vessel-days to be traded across countries (as is the case for the PNA). As before, a closure in Country 1 lowers the value of vessel-days in that country (because the fishable area shrinks). But increased biomass in other countries causes their vessel day prices to increase. As a result, vessel-days from Country 1 are traded to Countries 2 to 9, until prices equalize across countries (Fig. S3). Under this market design, revenue losses to Country 1 are less than 1\%, compared to the base case with no reserve (Fig. \ref{fig:PNA_model}B; note that the vertical axis now only ranges from 0 to 0.8\% instead of from 0 to 100\% as in Fig. \ref{fig:PNA_model}A). With trading, the relative revenue drop will always be smaller than the relative effort drop, and the opposite is observed when there is no trading (Fig. S4). Overall, this shows that 88\% to 99\% of the costs of conservation can be avoided if markets are designed to allow trading (Fig. \ref{fig:PNA_model}C).

@@

A large-scale MPA was recently implemented in PNA waters, providing the ideal empirical setting to test our predictions. In January 2015, Kiribati closed 11.5\% of its EEZ to implement the Phoenix Islands Protected Area (PIPA), effectively displacing all fishing effort within its boundaries \cite{mccauley_2016,mcdermott_2018}. By protecting important tuna spawning habitat, PIPA may provide important recruitment and biomass benefits to adjacent waters \cite{hernndez_2019}. We combine vessel-level tracking data \cite{kroodsma_2018} and country-level license revenue data reported by the Pacific Islands Forum Fisheries Agency (FFA) \cite{ffa_2017} to quantify the displacement of vessel-days and the likely costs of conservation. Of the 318 tuna purse seine vessels that fished in PNA waters between 2012 and 2018, 64 ``displaced'' vessels fished within PIPA at least once prior to its implementation and 254 ``non-displaced vessels'' never fished in PIPA waters but fished within PNA waters. We use the vessel-level tracking data to calculate vessel-days in the same way that the PNA does (see Methods for details) and to follow the redistribution of fishing effort before and after the implementation of PIPA.

Consistent with our model's prediction, after PIPA was implemented, displaced vessels relocated largely outside of Kiribati, and into other PNA countries' waters (Fig \ref{fig:fishing_raster_diff} and Figs. S6). Indeed, from 2014 to 2015, displaced vessels spent 2,115 fewer vessel-days (a 25\% reduction) in Kiribati, and 2,298 fewer vessel-days in PNA waters (a 17 \% reduction; Fig. \ref{fig:empirical}). On the other hand, non-displaced vessels spent 4,656 additional days in Kiribati during 2015, and an additional 9,598 vessel-days in PNA waters. By 2018, we observe a net decrease of vessel-days within Kiribati, from 12,671 in 2014 to just 7,677 in 2018, with displaced vessels driving the decrease (Fig. \ref{fig:empirical}A). However, aggregate effort at the PNA-level remains relatively constant and we do not observe a ``fishing the line'' effect (Fig S7). The reduction in effort in Kiribati and constant effort at the PNA-level suggest that trading facilitated redistribution of effort within PNA waters, just as the model predicts.

The decrease of vessel-days in Kiribati could be alternatively (or jointly) explained by changes in oceanographic conditions that drive the distribution of target species and fishing vessels. For example, as El Ni\~no events develop, tuna species are known to shift longitudinally across PNA waters, causing vessels to redistribute \cite{aqorau_2018,hanich2018unraveling}. However, the aggregate decrease that we observe for Kiribati in Fig. \ref{fig:empirical} is driven by the relocation of vessels that historically fished within PIPA, not by the entire fleet. Oceanographic conditions should be influencing the entire fleet whereas the closure of PIPA should have a stronger impact on vessels that used to fish there. Thus, the large drop in effort within Kiribati for the displaced vessels (relative to the non-displaced vessels) is consistent with the argument that PIPA displaced these vessels to waters outside of Kiribati. One could argue that differences in vessel characteristics between displaced and non-displaced vessels may influence their ability to redistribute or take advantage of different oceanographic conditions in Kiribati. It should be noted that, on average, displaced vessels have smaller crew sizes, more engine power, are larger than non-displaced vessels, fished more in the PNA in 2014, and are more likely to be registered to the Republic of Korea (Tables S2 and S3, Fig. S5). Thus, another alternative explanation for the observed changes in fishing patterns is a change in international relations between Kiribati (or other PNA countries) and the flag country of a particular vessel, although we are not aware of any occurring between 2012 and 2018. Ruling out alternative explanations for observed effects is an important component of careful empirical evaluation of MPAs \cite{ferraro2018causal}: as such, our results should not be interpreted as the direct causal impacts of PIPA but are better viewed as patterns that are consistent with the predictions of our theoretical model.

@@

%%%% PLEASE TAKE A LOOK AT THE SECOND SENTENCE OF THIS PARAGRAPH %%%%
While trading would allow Palau, Kiribati, and other PNA members to reduce revenue losses from large-scale conservation, our model indicates that moving from the current 40\% biomass-based allocation rule to a 100\% biomass-based rule would ensure long-term financial security in the presence of large-scale MPAs, and further incentivize conservation within the PNA. This is a model prediction that cannot be empirically tested in the PNA context because the allocation rule has not been experimentally modified. Nevertheless, the PNA countries have shown that rights-based management of transboundary resources can result in large management and economic benefits \cite{havice_2013,aqorau_2018}. By facilitating trade and allocating rights based on biomass, they may also have become pioneers in effective large-scale marine conservation in a market-based setting.

There are, however, a series of important considerations to take into account. First, in our model only one of the countries considers the implementation of a protected area. An interesting extension could consider cooperative conservation. In that setting, a group of countries could coordinate on an optimized, large-scale strategic closure -perhaps using vessel-day trading as a means of compensation- in a manner similar to the model above. Future research could explore the role of heterogeneous costs and benefits between countries, and the way in which these can shape conservation outcomes in a market-based setting. A second consideration is the role of the high seas: environmental markets require secure property rights, which are lacking in the high seas \cite{crespo_2019}. Benefits accruing to the high seas could potentially be eroded by the prevailing open access conditions. Therefore, a conservation-minded nation has no mechanism to capture the benefits provided to the high seas, highlighting the importance of empowering high seas governance and transboundary cooperation \cite{aqorau_2018,crespo_2019}. Finally, our work focuses on fisheries and large-scale conservation, but the framework could potentially be expanded and applied to other systems and natural resources. For example, similar mechanisms could be implemented in markets for tradable water rights \cite{rosegrant_1994} or game hunting \cite{hutton_2003} where secure property rights and the presence of a market may incentivize users to conserve the resource in question.

@@

Automatic Identification Systems (AIS) are on-board devices that provide at-sea safety and prevent ship collisions by broadcasting vessel position, course, and activity to surrounding vessels. These broadcast messages can be received by satellites and land-based antennas. We use AIS data provided by Global Fishing Watch \cite{kroodsma_2018} to track 318 tuna purse seiners that fished within the PNA. For every georeferenced position, we observe the time spent (defined as the time since the last position), and whether the vessel was actively fishing versus only transiting. Of the 318 tuna purse seine vessels that fished in PNA waters between 2012 and 2018, 64 ``displaced'' vessels fished within PIPA at least once prior to its implementation, the remianing 254 ``non-displaced vessels'' never fished in PIPA waters. Our dataset contains more than 37 million geo-referenced positions for these 318 tuna purse seiners. We use these data to calculate vessel-days (the metric used by the PNA), and to track the spatial redistribution of displaced vessels. A comparision of vessel characteristics between displaced and non-displaced vessels is presented in Tables S2-S3 and Figure S5.

We use these data to calculate the number of vessel-days that the 318 purse seiners spent in each PNA country and in PNA waters as a whole (Fig. \ref{fig:empirical}). The vessel-day equivalent of a day of fishing depends on vessel size, a measure used to control for effort creep. A day of activity by vessels smaller than 50 meters long counts as half a vessel-day, a day of activity by vessels 50 - 80 meters long counts as one day, and a day of activity by vessels larger than 80 meters counts as 1.5 vessel-days. Vessel length is an observable characteristic in our dataset, and therefore vessel-days calculated in our analyses correspond to the PNA definition of vessel-days.

@@

We begin by filtering the data to keep only positions labeled as fishing events. We then create a gridded version of the data for each year and group (\emph{i.e.} displaced and non-displaced) by binning the coordinates to a 1-degree grid and summing all fishing hours for a given grid cell, resulting in 14 (7 years for two groups) gridded datasets of fishing hours. For each group of vessels, we then calculate the average fishing hours before (2012 - 2014, inclusive) and after (2015 - 2018, inclusive) the implementation of PIPA, resulting in 4 datasets of mean fishing effort (before and after for displaced and non-displaced vessels). We then calculate the change in effort allocation between these two periods (after minus before) for each group (Fig. \ref{fig:fishing_raster_diff}A-B). The redistribution by non-displaced vessels (Fig. \ref{fig:fishing_raster_diff}B) therefore provides a baseline of redistribution. We then compare the changes of displaced vessels to those by non-displaced vessels (Fig. \ref{fig:fishing_raster_diff}C). The spatial redistribution patterns of displaced vessels relative to non-displaced vessels suggest that some relocated to other waters in Kiribati (\emph{i.e.} the Gilbert islands and Line islands), but also the Marshall Islands, Tuvalu, Nauru, and the high seas.

\subsubsection{Revenues}

### 2019-10-26 docs/Supplementary_materials.tex
Updating with John's changes

We observe four characteristics for every vessel: crew size, engine power, vessel length, and tonnage capacity. Figure \ref{fig:balance_density_plot} shows the distribution of these variables for each group of vessels. Table \ref{tab:balance_table} presentsibut the mean, and standard deviation of each observable. Group of vessels are similar on observables, and choice to fish inside PIPA before implementation does not appear to be driven by observable characteristics.

\caption{\label{tab:balance_table}Mean values on observable characteristics by vessel for displaced (n = 64), and non-displaced vessels (n = 254). Numbers in parentheses indicate standard deviation. The last column contains the difference in means (t-scores), with asterisks inicating significant differences as indicated by a two-tailed t-test (* p $<$ 0.1; ** p $<$ 0.05; *** p $<$ 0.01).}
\centering
\begin{tabular}{l|l|l|l}
\hline
Characteristic & Displaced & Non-displaced & Difference\\
\hline
Crew size (n) & 26.38 (3.94) & 30.46 (6.25) & 4.08 (6.49) ***\\
\hline
Engine Power (KW) & 2983.6 (558.76) & 2559.89 (588.28) & -423.71 (-5.36) ***\\
\hline
Length (m) & 74.23 (9.71) & 68.97 (8.42) & -5.25 (-3.97) ***\\
\hline
Tonnage (GT) & 1718.14 (653.38) & 1383.41 (533.56) & -334.73 (-3.79) ***\\
\hline
\end{tabular}
\end{table}

>>>>

We observe six characteristics for every vessel: flag, crew size, engine power, vessel length, tonnage capacity, and fishing hours within PNA waters in 2014. Figure \ref{fig:balance_density_plot} shows the distribution of the numeric variables for each group of vessels. Table \ref{tab:balance_table} presents the mean and standard deviation of each observable, and table \ref{tab:balance_table_flags} shows the composition of each group by flag. On average, displaced vessels have smaller crew sizes, more engine power, are larger than non-displaced vessels, and fished more in PNA waters during 2014. The largest relative difference is in terms of gross tonnage.

%% TABLE FOR NUMERIC VARIALBES
\begin{table}[H]
\caption{\label{tab:balance_table}Mean values on observable characteristics by vessel for displaced (n = 64), and non-displaced vessels (n = 254). Numbers in parentheses indicate standard deviation. The last column contains the difference in means (t-scores), with asterisks inicating significant differences as indicated by a two-tailed t-test (* p $<$ 0.1; ** p $<$ 0.05; *** p $<$ 0.01).}
\centering
\begin{tabular}{l|l|l|l}
\hline
Characteristic & Displaced & Non-displaced & Difference\\
\hline
Crew size (n) & 26.38 (3.94) & 30.46 (6.25) & 4.08 (6.49) ***\\
\hline
Engine Power (KW) & 2983.6 (558.76) & 2559.89 (588.28) & -423.71 (-5.36) ***\\
\hline
Length (m) & 74.23 (9.71) & 68.97 (8.42) & -5.25 (-3.97) ***\\
\hline
PNA fishing in 2014 (hours) & 667.57 (489.24) & 529.33 (380.11) & -138.24 (-1.89) *\\
\hline
Tonnage (GT) & 1718.14 (653.38) & 1383.41 (533.56) & -334.73 (-3.79) ***\\
\hline
\end{tabular}
\end{table}

%% BALANCE TABLE FOR FLAGS
\begin{table}[t]
\caption{\label{tab:balance_table_flags}Proportion of vessel flags by group. Note that we do not observe the flag for two vessels (0.78\%) in the non-displaced group.}
\centering
\begin{tabular}{l|r|r}
\hline
Flag & Non-displaced & Displaced\\
\hline
CHN & 10.24 & 0.00\\
\hline
ECU & 1.57 & 4.69\\
\hline
ESP & 0.39 & 4.69\\
\hline
FSM & 8.27 & 3.12\\
\hline
GTM & 0.39 & 1.56\\
\hline
JPN & 16.93 & 0.00\\
\hline
KIR & 2.76 & 12.50\\
\hline
KOR & 3.54 & 45.31\\
\hline
MEX & 1.18 & 0.00\\
\hline
MHL & 3.54 & 1.56\\
\hline
NIC & 0.39 & 0.00\\
\hline
NRU & 0.39 & 0.00\\
\hline
NZL & 1.18 & 3.12\\
\hline
PAN & 0.79 & 0.00\\
\hline
PHL & 7.87 & 1.56\\
\hline
PNG & 11.42 & 1.56\\
\hline
SLB & 1.57 & 0.00\\
\hline
SLV & 0.79 & 0.00\\
\hline
TWN & 11.81 & 3.12\\
\hline
USA & 12.20 & 17.19\\
\hline
VUT & 1.97 & 0.00\\
\hline
Not reported & 0.79 & 0.00\\
\hline
Total & 100.00 & 100.00\\
\hline
\end{tabular}
\end{table}

### 2019-10-26 docs/old_ms/Markets_Conservation_revised_Oct22.tex
Updating with John's changes

As predicted by our model, the implementation of PIPA resulted in a decrease in fishing effort within Kiribati's water without large revenue losses. Kiribati's reported revenue increased from US\$127.3 million in 2014 to US\$148.8 million in 2015, before decreasing to US\$118.3 million in 2016 (Fig. \ref{fig:empirical}C). The increase and subsequent decrease in revenues matches the vessel-day patterns observed for Kiribati from 2014 to 2016 (Fig. \ref{fig:empirical}D). But, critically, the drop in revenue in 2016 (20\%) is smaller than the drop in VDS effort (35\%). This supports a key prediction of our model: with trading, the relative revenue drop will always be smaller than the relative effort drop but, without trading, the opposite relationship would hold (Fig. S4). At the PNA level, total revenues showed a net increase of \$64.7 and \$28 million USD for 2015 and 2016 respectively (Fig. S8), despite the PIPA closure.

>>>>

As predicted by our model, the implementation of PIPA resulted in a decrease in fishing effort within Kiribati's water without large revenue losses. Kiribati's reported revenue increased from US\$127.3 million in 2014 to US\$148.8 million in 2015, before decreasing to US\$118.3 million in 2016 (Fig. \ref{fig:empirical}C). The increase and subsequent decrease in revenues matches the vessel-day patterns observed for Kiribati from 2014 to 2016 (Fig. \ref{fig:empirical}D). But, critically, the drop in revenue in 2016 (20\%) is smaller than the drop in VDS effort (35\%). This supports a key prediction of our model: with trading, the relative revenue drop will be smaller than the relative effort drop but, without trading, the opposite relationship would hold (Fig. S4). At the PNA level, total revenues showed a net increase of \$64.7 and \$28 million USD for 2015 and 2016 respectively (Fig. S8), despite the PIPA closure.
